---
title: "Final Presentation"
author: "Amber Hu, Melody He, Elaine Liu, Allen Dong"
date: '2022-06-22'
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(tidyverse)
library(forcats)
library(gridExtra)
library(RSQLite)
library(jsonlite)
library(gtable)
library(grid)
library(latex2exp)
library(gridBase)
library(nnet)
library(magrittr)
library(ggplot2)
library(data.table)
library(ggmap)
library(ggalluvial)
library(stringr)
library(ggpubr)
library(gridExtra)
library(psych)
library(gt)
library(rgdal)
library(mapproj)
library(tidyr)

```

## Shiny Presentation

This R Markdown presentation is made interactive using Shiny. The viewers of the presentation can change the assumptions underlying what's presented and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Interactive Plot

```{r}

```

## Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Location Alluvial Plot

```{r}
full_data <- fread("/Users/allendong/Desktop/chicagocrimes.csv")

# subsetting data
data1 <- full_data
names(data1)[names(data1) == 'Primary Type'] <- 'type'
names(data1)[names(data1) == 'Location Description'] <- 'location'

data1$location <- str_replace_all(data1$location,
                                  "PARKING LOT/GARAGE\\(NON\\.RESID\\.\\)", 
                                  "PARKING LOT/GARAGE")

top_type <- data1 %>% 
  count(type) %>% 
  top_n(8, n) %>% 
  pull(type)

top_location <- data1 %>% 
  filter(type %in% top_type) %>% 
  count(location) %>% 
  top_n(6, n) %>% 
  pull(location)

filtered_data1 <- data1 %>% 
  filter(type %in% top_type & location %in% top_location)

grouped_data1 <- filtered_data1 %>% group_by(type, location) %>% summarise(Freq = n())  
personal_theme = theme(plot.title = element_text(hjust = 0.5))

ggplot(data = grouped_data1,
       aes(axis1 = type, axis2 = location, y = Freq)) +
  geom_alluvium(aes(fill = location)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c('Type of Crime', 'Location of Crime'),
                   expand = c(0.20, 0.05)) +
  scale_fill_viridis_d() +
  theme_void() +
  theme(legend.position = "none") + 
  ggtitle("Summary of Crime Cases in Chicago \nMapping Major Types of Crime to Major Locations") +
  personal_theme

## Source: https://data.cityofchicago.org/Public-Safety/Boundaries-Police-Districts-current-/fthy-xz3r
df <- fread("/Users/allendong/Desktop/chicagocrimes.csv")
df <- select(df, -ID, -`Case Number`, -IUCR, -Block, -Description, -`Updated On`, -Ward, -Location)

# import boundaries of police districts
bounds <- readOGR("/Users/allendong/Desktop/districts.shp")
# convert to dataframe
bounds_df <- fortify(bounds)
bounds_df$id <- as.factor(bounds_df$id)

## map labels for district numbers to id in boundaries dataframe
labs <- data.frame(bounds$dist_num)
labs <- mutate(labs, id = strtoi(rownames(labs)) - 1)
colnames(labs) <- c("District", "id")
bounds_df <- merge(bounds_df, labs, by = "id")
labs <- select(labs, District)

# centroid calculations
centroids <- as.data.frame(coordinates(bounds))
colnames(centroids) = c("c.long", "c.lat")

## district counts table
district_counts <- as.data.frame(table(df$District))
colnames(district_counts) = c("District", "Frequency")
district_counts$Frequency <- district_counts$Frequency/1000

dis.map <- merge(bounds_df, district_counts, by = 'District', all = TRUE)
labs.df <- cbind(labs, centroids)
labs.df <- labs.df[order(as.numeric(labs.df$District)), ]
## manually adding centroid coordinates for 16th district
labs.df[15, ] <- c(16, -87.91, 41.98)
labs.df[nrow(labs.df)+1, ] <- c(16, -87.77, 41.97)
## convert longitude and latitudes to doubles
labs.df$c.long <- as.double(labs.df$c.long)
labs.df$c.lat <- as.double(labs.df$c.lat)

dis.map <- merge(dis.map, labs.df, by = 'District', all = TRUE)
dis.map <- arrange(dis.map, group, order)
```

## Title
```{r, echo = FALSE, warning = FALSE}
map <- ggplot(dis.map[dis.map$District != 31,], aes(x = long, y = lat, group = group, fill = Frequency)) +
  coord_map("polyconic") + 
  geom_polygon(colour = "black")  +
  ggtitle("Frequency of Crime by Police District") + 
  scale_fill_gradient(low="white", high="slateblue3") +
  geom_text(aes(label = `District`, x = c.long, y = c.lat, family = 'serif'), size = 3) + xlab("longitude") + ylab("latitude") + theme(text = element_text(family = "serif"))
map <- map + labs(fill = "Frequency (in thousands)")
map
```


```{r, include=FALSE}
# subset every seventh row
ggmap::register_google(key = "API KEY")

LatLonCounts <- as.data.frame(table(round(data1$Longitude,2), round(data1$Latitude,2)))
LatLonCounts$Long <- as.numeric(as.character(LatLonCounts$Var1))
LatLonCounts$Lat <- as.numeric(as.character(LatLonCounts$Var2))
LatLonCounts2 <- subset(LatLonCounts, Freq > 0)
p <- ggmap(get_googlemap(center = c(lon = -87.6298, lat = 41.8781), zoom = 11, scale = 2, maptype ='terrain', color = 'color', alpha = 0.5))
```

## Location Plot

```{r, warning=FALSE, message = FALSE}
p + geom_tile(data = LatLonCounts2, aes(x = Long, y = Lat, alpha = Freq), fill = "red")
```

## Location plot 2
```{r, warning = FALSE, fig.width = 8, fig.height = 5, message = FALSE}
sub <- data1[which(1:nrow(data1) %% 56 == 0),]
p + geom_point(aes(x = Longitude, y = Latitude, col = Arrest), data = sub, alpha = 0.01, size = 3)
```

```{r, include=FALSE}
data6 <- full_data
names(data6)[names(data6) == 'Primary Type'] <- 'type'

top_type <- data6 %>% 
  count(type) %>% 
  top_n(10, n) %>% 
  pull(type)

grouped_data6 <- data6 %>% 
  filter(type %in% top_type) %>% 
  group_by(type, Arrest) %>%
  summarise(Freq = n())   
```

##

```{r, warning=FALSE, message = FALSE}
ggplot(data = grouped_data6, aes(fill=Arrest, y=Freq, x=reorder(type, Freq, sum))) + 
  geom_bar(position="stack", stat="identity") + 
  ggtitle("Crime Arrest By Primary Type") +
  scale_fill_viridis_d() + 
  labs(x = 'Primary Type of Crime', y = 'Frequency of Crime') +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1))
```

```{r, include=FALSE}
convert_time <- function(date) {
  hour <- as.numeric(str_sub(date, 12, 13))
  is.am <- (str_sub(date, -2, -1) == 'AM') 
  ifelse(hour == 12, ifelse(is.am, 0, 12), ifelse(is.am, hour, hour + 12))
}

data4 <- full_data %>%
  mutate(Time = convert_time(Date)) %>%
  group_by(Time, Arrest) %>% 
  summarise(Freq = n())  

```

## Title

```{r, warning=FALSE, message = FALSE}
ggplot(data = data4, aes(fill=Arrest, y=Freq, x=Time)) + 
  geom_bar(position="fill", stat="identity") + 
  ggtitle("Proportion of Crime Arrested By Hour of the Day") +
  scale_x_continuous(breaks = round(seq(0, 23, by = 1),1)) +
  scale_fill_viridis_d() +
  coord_polar() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r, include=FALSE}
data5 <- full_data %>% 
  group_by(Year, Arrest) %>%
  # filter(Year < 2022) %>% 
  summarise(Freq = n())  
```

## Title

```{r, warning=FALSE, message = FALSE}
ggplot(data = data5, aes(fill=Arrest, y=Freq, x=Year)) + 
  geom_bar(position="fill", stat="identity") + 
  ggtitle("Proportion of Crime Arrested By Year") +
  scale_y_continuous(breaks = round(seq(0, 1, by = 0.10),1)) +
  scale_fill_viridis_d()

df <- fread("/Users/allendong/Desktop/chicagocrimes.csv")
# df <- select(df, -ID, -`Case Number`, -IUCR, -Block, -Description, -`Updated On`, -Ward, -Location)
## secondary dataset
data1 <- fread("/Users/allendong/Desktop/poverty.csv")
data1 <- data1[1:77,]
```

## Title

```{r, warning = FALSE, fig.width = 6, fig.height = 6, message = FALSE}
library(psych)
g1 <- ggplot(data = data1, aes(x = `PERCENT HOUSEHOLDS BELOW POVERTY`)) + 
  geom_histogram(binwidth = 10, color = "plum", fill = "plum1") + xlab("% Households Below Poverty") + ylab("Frequency")
g2 <- ggplot(data = data1, aes(x = `PERCENT AGED 16+ UNEMPLOYED`)) + 
  geom_histogram(binwidth = 10, color = "paleturquoise", fill = "paleturquoise1") + xlab("Percent Unemployed") + ylab("Frequency")
g3 <- ggplot(data = data1, aes(x = `PER CAPITA INCOME`)) + 
  geom_histogram(binwidth = 10000, color = "lightpink", fill = "lightpink1") + xlab("Per Capita Income") + ylab("Frequency")
g4 <- ggplot(data = data1, aes(x = `HARDSHIP INDEX`)) + 
  geom_histogram(binwidth = 20, color = "plum", fill = "plum1") + xlab("Hardship Index (1-100)") + ylab("Frequency")
grid.arrange(g1, g2, g3, g4, nrow = 2)
```

##
```{r, echo = FALSE}
stats <- sapply(data1[,c('PERCENT HOUSEHOLDS BELOW POVERTY', 'PERCENT AGED 16+ UNEMPLOYED', 'PER CAPITA INCOME', 'HARDSHIP INDEX')], summary)
stats <- as.data.frame(round(t(stats), 2))
row.names(stats) <- c("% Households Below Poverty", 
                      "% Aged 16+ Unemployed",
                      "Per Capita Income ($)", 
                      "Hardship Index (1-100)")
stats %>%
  gt(rownames_to_stub = TRUE) %>%
  tab_header(title = "Socioeconomic Indicators Summary (2008-2012)")
```

```{r, warning=FALSE, message = FALSE, include = FALSE}

## obtain subset of Year = 2010
sub <- df[df$Year == 2010,]

## frequency table dataframe
comm <- data.frame(table(sub$`Community Area`)) 
colnames(comm) <- c("Community Area","Crime Rate")
## remove community area 0
comm <-comm[comm$`Community Area` != 0,]
# find crime rate per 10,000 people
#comm$`Crime Rate` <- comm$`Crime Rate`/(as.numeric(data1$POPULATION[-78])) * 10000

## merge frequency dataframe with secondary dataset
comm <- merge(comm, data1, by.x = "Community Area", by.y = "Community Area Number")
colors <- c("red", "orange", "green", "blue", "magenta", "purple")

## correlation plots with correlation coefficient and p-values
p1 <- ggplot(data = comm, aes(x = `Crime Rate`, y = `PERCENT HOUSEHOLDS BELOW POVERTY`)) + 
  geom_point(col = colors[1]) + 
  geom_smooth(method='lm', formula= y~x, size = 0.25, col = "black", se = FALSE) + 
  xlab("Crime Rate (per 10000 people)")+ ylab("% Households Below Poverty") +
  stat_cor(label.x = 600, label.y = 55)

p2 <- ggplot(data = comm, aes(x = `Crime Rate`, y = `PERCENT AGED 16+ UNEMPLOYED`)) + 
  geom_point(col = colors[2]) + 
  geom_smooth(method='lm', formula= y~x, size = 0.25, col = "black", se = FALSE) + 
  xlab("Crime Rate (per 10000 people)") + ylab("% unemployed (Ages 16+)") +
  stat_cor(label.x = 600, label.y = 55)

p3 <- ggplot(data = comm, aes(x = `Crime Rate`, y = `PER CAPITA INCOME`, )) + 
  geom_point(col = colors[4]) +geom_smooth(method='lm', formula= y~x, size = 0.25, col = "black", se = FALSE) + 
  xlab("Crime Rate (per 10000 people)") +
  ylab("Per Capita Income") + 
  stat_cor(label.x = 500, label.y = 80000)

p4 <- ggplot(data = comm, aes(x = `Crime Rate`, y = `HARDSHIP INDEX`)) + 
  geom_point( col = colors[5]) + 
  geom_smooth(method='lm', formula= y~x, size = 0.25, col = "black", se = FALSE) + 
  xlab("Crime Rate (per 10000 people)") + ylab("Hardship Index (0-100)") + 
  stat_cor(label.x = 600, label.y = 100)
```

## Title
```{r, echo = FALSE}
grid.arrange(p1, p2, p3, p4, nrow = 2, top="Four Socioeconomic Factors vs. Crime Rate, 2010")
```


